Clines from PDE

### 2015-10-26 notes/diffusion-local-times.Rmd
.

Now let's get local times for those walks.
```{r lineage_local_times,cache=TRUE,depends="sim_lineage"}
xx <- seq( min(rws), max(rws), length.out=200 )
xmids <- xx[-1]-diff(xx)/2
lts <- attr(rws,"dt")*apply( matrix(findInterval(rws,xx),nrow=nrow(rws)), 2, tabulate, nbins=length(xx)-1 )
layout( t(1:2), widths=c(3,1) )
matplot(lts[,1:30],xmids,type='l',xlab="local time", ylab="space")
plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
```

Here is the distribution of $T=\int_0^{`r tau`} f_H(X_t) dt$:
```{r dist_het_time, cache=TRUE, depends="sim_het_time"}
het.time <- weighted_time( rws, tanh_het, s=s )
layout(1:2)
hist( het.time, breaks=50, xlab="time in hets", freq=FALSE, main="histogram of T" ) 
hist( log(het.time), breaks=50, xlab="log(time in hets)", freq=FALSE , main="histogram of log(T)") 
```

@@

First, two quick checks: First, by setting $f_H=1$,
we should have the solution $u(x,t,r) = \exp(-tr)$.
Then, by setting $\sigma$ to be very small, the solution at $x$ should be $\exp(-trf_H(x))$.
```{r solve_pde_check, cache=TRUE, depends="solve_pde_setup"}
layout(t(1:2))
test.hetfn <- setup.prop.1D(grid=thegrid, func=function(x){rep(1,length(x))})
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * test.hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
matplot(pde.soln,type='l',ylab="probability of no recomb", col='black',log='y')
lines(tt,exp(-r*tt),lwd=2,lty=2,col='red')
## no diffusion
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1e-6, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
plot.x <- sort(match(sample(thegrid$x.mid[thegrid$x.mid>-5 & thegrid$x.mid<5],12),thegrid$x.mid))
matplot(pde.soln[,plot.x],type='l',ylab="probability of no recomb", col=rainbow(length(plot.x)), lty=1 )
matpoints( exp(outer( tt, -r*hetfn$mid[plot.x], "*" )), lty=4, col=rainbow(length(plot.x)), pch=20 )
```
**Looks good.**

Now, to solve the real problem:
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
image(pde.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

It's reassuring the solution is between 0 and 1.
Let's spot check that with the simulation.
Here it is at $t=`r tau`$:
```{r check_pde_soln, cache=TRUE}
check.t <- function (alpha) {
    t.tau <- which.min(abs(tt-alpha*tau))
    probs <- exp( - weighted_time( rws, tanh_het, s=s, use.t=1:floor(nrow(rws)*alpha) ) )
    plot(rws[1,],probs,xlab='initial location', ylab=paste('probability of no recomb by t=',alpha*tau),ylim=c(0,1))
    lines(lowess(rws[1,],probs,f=0.05),col='red',lwd=2)
    lines(thegrid$x.int,pde.soln[t.tau,],col='green',lwd=2)
    legend("topright",lwd=2,col=c("red","green"),legend=c("mean","numerical solution"))
}
check.t(1)
```
and, at other values of $t$:
```{r check_pde_soln_2, cache=TRUE, depends="check_pde_soln"}
layout(t(1:2))
check.t(0.1)
check.t(0.2)
check.t(0.4)
check.t(0.8)
```
**Looks pretty good.**

@@

Actually, the cline gets set up
in a time-inhomogeneous way.
Let's do that, by numerical solution,
then run the lineage back through it.
The forwards-time evolution of the cline,
with isotropic density and migration,
solves the PDE (Barton 1979):
$$\begin{aligned}
  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 p(x,t)+ s p(x,t)(1-p(x,t))(2p(x,t)-1) .
\end{aligned}$$
```{r fwds_pde, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
yinit <- ifelse( thegrid$x.mid>0, 0, 1 )
tt <- seq(0,tau,length.out=100)
fwds.pde <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, D=1/2, dx=thegrid)$dC
    list( tran + s * y * (1-y) * (2*y-1) )
}
fwds.soln <- ode.1D( y=yinit, times=tt, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
image(fwds.soln,legend=TRUE,main="mean cline development", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
Now, let's run the killed lineage backwards in it:
```{r fwds_back_pde, cache=TRUE, depends="fwds_pde"}
yinit <- rep(1.0,thegrid$N)
rev.pde <- function (t,y,parms,...) {
    p <- cline_interp(tau-t,soln=fwds.soln)
    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline_interp(tau-t)), interp="linear")
    tran <- tran.1D(C=y, A=p, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * (1-p) * y )
}
rev.soln <- ode.1D( y=yinit, times=tt, func=rev.pde, nspec=1, tcrit=tau ) # note FIRST COLUMN IS TIME
image(rev.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

Let's compare the case where the cline is set up instantly or not:
```{r compare_rightness, cache=TRUE,depends=c("fwds_back_pde","solve_pde_setup","setup_fns")}
layout(t(1:2))
fb.p <- clineplot(rev.soln,main="forwards-backwards")
st.p <- clineplot(pde.soln,main="stationary cline")
layout(1)
plot(fb.p,st.p,asp=1,pch=20,cex=0.5,xlab="forwards-backwards",ylab="stationary",col=rainbow(ncol(fb.p))[col(fb.p)])
abline(0,1,lwd=3)
```

@@

Of course, we aren't actually trying to estimate the distribution of $T$ itself,
but rather the probabilities of recombination (the Laplace transform).
Here are these probabilities, as a function of $r$
(the second plots are on a log-log scale):
```{r recomb_prob_comparisons, cache=TRUE, depends="sim_general"}
rvals <- exp(seq(log(.001),log(0.5),length.out=400))
precombs <- lapply( lt.comparison, function (lts) {
        sapply( lts, function (lt) sapply( rvals, function (r) { mean((-1)*expm1(-r*lt)) } ) )
    } )
layout(t(1:2))
for (k in seq_along(precombs)) {
    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1 )
    legend("topright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
}
for (k in seq_along(precombs)) {
    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), log='xy', main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1  )
    legend("bottomright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
}
```

@@

Now, let's check on the local time formula.
The number of steps the random walk spends at position $k$ on a grid with spacing $\epsilon$
and a time step of \gamma
should be $\epsilon \times \ell(k\epsilon) / \gamma$.
First, the local time at 0:
```{r local_times_0, cache=TRUE, depends="sim_rw"}
xh <- hist(colSums(abs(rws)<=1.5*eps)/ntimes/(3*eps),main="local time at 0", breaks=30)
lines( xh$mids, sum(xh$counts)*diff( 2*pnorm(xh$breaks)-1 ), col='red' )
```
What about the local time at other locations?
Note there is an atom at 0 now
(that is depicted by the dots in the figures below),
and that the discretization could trip us up;
the colors (red/green/blue) use the (mid/upper/lower) points of the interval that we count in.
```{r local_times, cache=TRUE, depends="sim_rw"}
layout(matrix(1:4,nrow=2))
for (loc in c(0.1, 0.2, 0.4, 0.6)) {
    lts <- colSums(abs(rws-loc)<=1.5*eps)/ntimes/(3*eps)
    zeros <- (lts==0)
    pred.zero <- 2*pnorm(abs(loc)+c(0,1.5*eps,-1.5*eps))-1
    xh <- hist(lts[!zeros],main=paste("local time at",loc), breaks=30, ylim=range(c(0,pred.zero*nwalks)))
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc)+xh$breaks)-1 ), col='red' )
    points(rep(0,4), c(sum(zeros),pred.zero*nwalks), col=1:4, pch=20, cex=2)
    # lower bound
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc+1.5*eps)+xh$breaks)-1 ), col=3 )
    # upper bound
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc-1.5*eps)+xh$breaks)-1 ), col=4 )
}
```
Clearly, the discretization is still tripping us up a bit,
but it's pretty good.

@@

Let's check that, too: why not?
```{r killing, cache=TRUE, depends="sim_rw"}
zeta.reflected
loc <- 0.1
compare.rates <- as.data.frame( t( sapply( c(0.25,0.5,1,1.5,2,2.5,3), function (rate) {
        visits <- colSums(abs(rws-loc)<=1.5*eps)
        killed <- ( rbinom(ncol(rws),size=visits,prob=rate/ntimes/(3*eps)) == 0 )
        c( 
          rate=rate,
          simulation=mean(killed), 
          theory=zeta.reflected(loc,rate),
          SE=sqrt(mean(killed)*(1-mean(killed))/nwalks)
          )
    } ) ) )
with(compare.rates, {
         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=c(0,1) ) 
         points( rate, simulation, pch=20 )
         arrows( x0=rate, y0=simulation-2*SE, y1=simulation+2*SE, angle=90, code=3 )
    } )
```

@@

Here are the two compared,
by smoothing the probabilities for the simulation:
```{r sim_clines, cache=TRUE, depends=c("sim_pushed","pushed_pde")}
prob.no.recomb <- sapply( rvals, function (r) { fitted(loess( exp(-r*sim.wt) ~ xinit, span=0.1)) } )
matplot(xinit,prob.no.recomb,type='l',ylab="prob of no recomb",
        xlab="space",col=rainbow(length(rvals)),lty=1,main=sprintf("t=%d",tau))
legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals,digits=2)))
for (k in seq_along(rvals)) {
    lines( thegrid$x.mid, pde.solns[[k]][findInterval(tau,pde.solns[[k]][,1]),-1], 
          lty=3, lwd=3, col=rainbow(length(rvals))[k] )
}
```
Well, that looks reasonable, but hard to tell if the discrepancy is due to the loess or not.

And, finally, here's some actual clines predicted by this model:
```{r pushed_cline,cache=TRUE,depends="pushed_pde"}
tau <- 100
tt <- seq(0,tau,length.out=100)
pde.solns <- lapply( rvals, function (r) {
        pde.fn <- function (t,y,parms,...) {
            tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
            if (any(is.na(tran))) { browser() }
            list( tran - r * hetfn$mid * y )
        }
        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
      } )
plot.tvals <- floor(seq(0,tau,length.out=7))[-1]
layout(t(1:2))
for (tval in plot.tvals) {
    matplot( thegrid$x.mid, 
            pcline(thegrid$x.mid, p=sapply( pde.solns, cline_interp, T=tval ) ), 
                   type='l', main=sprintf("t=%d",tval) 
        )
    legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals/2,digits=2)))
}
```

>>>>

Now let's get local times for those walks.
```{r lineage_local_times,cache=TRUE,dependson="sim_lineage"}
xx <- seq( min(rws), max(rws), length.out=200 )
xmids <- xx[-1]-diff(xx)/2
lts <- attr(rws,"dt")*apply( matrix(findInterval(rws,xx),nrow=nrow(rws)), 2, tabulate, nbins=length(xx)-1 )
layout( t(1:2), widths=c(3,1) )
matplot(lts[,1:30],xmids,type='l',xlab="local time", ylab="space")
plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
```

Here is the distribution of $T=\int_0^{`r tau`} f_H(X_t) dt$:
```{r dist_het_time, cache=TRUE, dependson="sim_het_time"}
het.time <- weighted_time( rws, tanh_het, s=s )
layout(1:2)
hist( het.time, breaks=50, xlab="time in hets", freq=FALSE, main="histogram of T" ) 
hist( log(het.time), breaks=50, xlab="log(time in hets)", freq=FALSE , main="histogram of log(T)") 
```

@@

First, two quick checks: First, by setting $f_H=1$,
we should have the solution $u(x,t,r) = \exp(-tr)$.
Then, by setting $\sigma$ to be very small, the solution at $x$ should be $\exp(-trf_H(x))$.
```{r solve_pde_check, cache=TRUE, dependson="solve_pde_setup"}
layout(t(1:2))
test.hetfn <- setup.prop.1D(grid=thegrid, func=function(x){rep(1,length(x))})
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * test.hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
matplot(pde.soln,type='l',ylab="probability of no recomb", col='black',log='y')
lines(tt,exp(-r*tt),lwd=2,lty=2,col='red')
## no diffusion
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1e-6, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
plot.x <- sort(match(sample(thegrid$x.mid[thegrid$x.mid>-5 & thegrid$x.mid<5],12),thegrid$x.mid))
matplot(pde.soln[,plot.x],type='l',ylab="probability of no recomb", col=rainbow(length(plot.x)), lty=1 )
matpoints( exp(outer( tt, -r*hetfn$mid[plot.x], "*" )), lty=4, col=rainbow(length(plot.x)), pch=20 )
```
**Looks good.**

Now, to solve the real problem:
```{r solve_pde, cache=TRUE, dependson="solve_pde_setup"}
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
image(pde.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

It's reassuring the solution is between 0 and 1.
Let's spot check that with the simulation.
Here it is at $t=`r tau`$:
```{r check_pde_soln, cache=TRUE}
check.t <- function (alpha) {
    t.tau <- which.min(abs(tt-alpha*tau))
    probs <- exp( - weighted_time( rws, tanh_het, s=s, use.t=1:floor(nrow(rws)*alpha) ) )
    plot(rws[1,],probs,xlab='initial location', ylab=paste('probability of no recomb by t=',alpha*tau),ylim=c(0,1))
    lines(lowess(rws[1,],probs,f=0.05),col='red',lwd=2)
    lines(thegrid$x.int,pde.soln[t.tau,],col='green',lwd=2)
    legend("topright",lwd=2,col=c("red","green"),legend=c("mean","numerical solution"))
}
check.t(1)
```
and, at other values of $t$:
```{r check_pde_soln_2, cache=TRUE, dependson="check_pde_soln"}
layout(t(1:2))
check.t(0.1)
check.t(0.2)
check.t(0.4)
check.t(0.8)
```
**Looks pretty good.**

@@

Actually, the cline gets set up
in a time-inhomogeneous way.
Let's do that, by numerical solution,
then run the lineage back through it.
The forwards-time evolution of the cline,
with isotropic density and migration,
solves the PDE (Barton 1979):
$$\begin{aligned}
  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 p(x,t)+ s p(x,t)(1-p(x,t))(2p(x,t)-1) .
\end{aligned}$$
```{r fwds_pde, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
yinit <- ifelse( thegrid$x.mid>0, 0, 1 )
tt <- seq(0,tau,length.out=100)
fwds.pde <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, D=1/2, dx=thegrid)$dC
    list( tran + s * y * (1-y) * (2*y-1) )
}
fwds.soln <- ode.1D( y=yinit, times=tt, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
image(fwds.soln,legend=TRUE,main="mean cline development", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
Now, let's run the killed lineage backwards in it:
```{r fwds_back_pde, cache=TRUE, dependson="fwds_pde"}
yinit <- rep(1.0,thegrid$N)
rev.pde <- function (t,y,parms,...) {
    p <- cline_interp(tau-t,soln=fwds.soln)
    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline_interp(tau-t)), interp="linear")
    tran <- tran.1D(C=y, A=p, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * (1-p) * y )
}
rev.soln <- ode.1D( y=yinit, times=tt, func=rev.pde, nspec=1, tcrit=tau ) # note FIRST COLUMN IS TIME
image(rev.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

Let's compare the case where the cline is set up instantly or not:
```{r compare_rightness, cache=TRUE,dependson=c("fwds_back_pde","solve_pde_setup","setup_fns")}
layout(t(1:2))
fb.p <- clineplot(rev.soln,main="forwards-backwards")
st.p <- clineplot(pde.soln,main="stationary cline")
layout(1)
plot(fb.p,st.p,asp=1,pch=20,cex=0.5,xlab="forwards-backwards",ylab="stationary",col=rainbow(ncol(fb.p))[col(fb.p)])
abline(0,1,lwd=3)
```

@@

Of course, we aren't actually trying to estimate the distribution of $T$ itself,
but rather the probabilities of recombination (the Laplace transform).
Here are these probabilities, as a function of $r$
(the second plots are on a log-log scale):
```{r recomb_prob_comparisons, cache=TRUE, dependson="sim_general"}
rvals <- exp(seq(log(.001),log(0.5),length.out=400))
precombs <- lapply( lt.comparison, function (lts) {
        sapply( lts, function (lt) sapply( rvals, function (r) { mean((-1)*expm1(-r*lt)) } ) )
    } )
layout(t(1:2))
for (k in seq_along(precombs)) {
    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1 )
    legend("topright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
}
for (k in seq_along(precombs)) {
    matplot( rvals, precombs[[k]], type='l', col=1:ncol(precombs[[k]]), log='xy', main=paste("x =",initvals[k]), ylim=range(precombs), ylab="prob of recomb", xlab='r', lty=1  )
    legend("bottomright",lty=1,col=1:ncol(precombs[[k]]),legend=colnames(precombs[[k]]))
}
```

@@

Now, let's check on the local time formula.
The number of steps the random walk spends at position $k$ on a grid with spacing $\epsilon$
and a time step of \gamma
should be $\epsilon \times \ell(k\epsilon) / \gamma$.
First, the local time at 0:
```{r local_times_0, cache=TRUE, dependson="sim_rw"}
xh <- hist(colSums(abs(rws)<=1.5*eps)/ntimes/(3*eps),main="local time at 0", breaks=30)
lines( xh$mids, sum(xh$counts)*diff( 2*pnorm(xh$breaks)-1 ), col='red' )
```
What about the local time at other locations?
Note there is an atom at 0 now
(that is depicted by the dots in the figures below),
and that the discretization could trip us up;
the colors (red/green/blue) use the (mid/upper/lower) points of the interval that we count in.
```{r local_times, cache=TRUE, dependson="sim_rw"}
layout(matrix(1:4,nrow=2))
for (loc in c(0.1, 0.2, 0.4, 0.6)) {
    lts <- colSums(abs(rws-loc)<=1.5*eps)/ntimes/(3*eps)
    zeros <- (lts==0)
    pred.zero <- 2*pnorm(abs(loc)+c(0,1.5*eps,-1.5*eps))-1
    xh <- hist(lts[!zeros],main=paste("local time at",loc), breaks=30, ylim=range(c(0,pred.zero*nwalks)))
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc)+xh$breaks)-1 ), col='red' )
    points(rep(0,4), c(sum(zeros),pred.zero*nwalks), col=1:4, pch=20, cex=2)
    # lower bound
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc+1.5*eps)+xh$breaks)-1 ), col=3 )
    # upper bound
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc-1.5*eps)+xh$breaks)-1 ), col=4 )
}
```
Clearly, the discretization is still tripping us up a bit,
but it's pretty good.

@@

Let's check that, too: why not?
```{r killing, cache=TRUE, dependson="sim_rw"}
zeta.reflected
loc <- 0.1
compare.rates <- as.data.frame( t( sapply( c(0.25,0.5,1,1.5,2,2.5,3), function (rate) {
        visits <- colSums(abs(rws-loc)<=1.5*eps)
        killed <- ( rbinom(ncol(rws),size=visits,prob=rate/ntimes/(3*eps)) == 0 )
        c( 
          rate=rate,
          simulation=mean(killed), 
          theory=zeta.reflected(loc,rate),
          SE=sqrt(mean(killed)*(1-mean(killed))/nwalks)
          )
    } ) ) )
with(compare.rates, {
         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=c(0,1) ) 
         points( rate, simulation, pch=20 )
         arrows( x0=rate, y0=simulation-2*SE, y1=simulation+2*SE, angle=90, code=3 )
    } )
```

@@

Here are the two compared,
by smoothing the probabilities for the simulation:
```{r sim_clines, cache=TRUE, dependson=c("sim_pushed","pushed_pde")}
prob.no.recomb <- sapply( rvals, function (r) { fitted(loess( exp(-r*sim.wt) ~ xinit, span=0.1)) } )
matplot(xinit,prob.no.recomb,type='l',ylab="prob of no recomb",
        xlab="space",col=rainbow(length(rvals)),lty=1,main=sprintf("t=%d",tau))
legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals,digits=2)))
for (k in seq_along(rvals)) {
    lines( thegrid$x.mid, pde.solns[[k]][findInterval(tau,pde.solns[[k]][,1]),-1], 
          lty=3, lwd=3, col=rainbow(length(rvals))[k] )
}
```
Well, that looks reasonable, but hard to tell if the discrepancy is due to the loess or not.

And, finally, here's some actual clines predicted by this model:
```{r pushed_cline,cache=TRUE,dependson="pushed_pde"}
tau <- 100
tt <- seq(0,tau,length.out=100)
pde.solns <- lapply( rvals, function (r) {
        pde.fn <- function (t,y,parms,...) {
            tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
            if (any(is.na(tran))) { browser() }
            list( tran - r * hetfn$mid * y )
        }
        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
      } )
plot.tvals <- floor(seq(0,tau,length.out=7))[-1]
layout(t(1:2))
for (tval in plot.tvals) {
    matplot( thegrid$x.mid, 
            pcline(thegrid$x.mid, p=sapply( pde.solns, cline_interp, T=tval ) ), 
                   type='l', main=sprintf("t=%d",tval) 
        )
    legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals/2,digits=2)))
}
```

### 2015-10-26 notes/pde-clines.Rmd
.

Let's solve this numerically.
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
r <- 0.1
s <- 0.1
sigma <- 5
zone.age <- 200  # this is tau
tt <- seq(0,zone.age,length.out=100)
pde.soln <- solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
pde.cline <- cline_from_soln(pde.soln,grid=thegrid)
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r),yaxt='n')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
# same thing, plotted with lines
clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
```
**Note:** the range should be wide enough these are *going to zero*
at the ends, or else we have to worry about the boundary conditions.

@@

We'll setup, run, and summarize simulations
as described in [this file](../sims/run-sims.html).
```{r setup_sims, cache=TRUE}
source("../sims/sim-fns.R")
zone.age    = 50
ndemes 		= 500
deme.size 	= 5
ninds 		= ndemes*deme.size
qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s, pos=c(0.5)))
# spatial locations
xx <- (1:ndemes)-0.5-ndemes/2
# run the simulation
sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=sigma,quiet=TRUE)
# post-process
sim.sums <- spBreaks(sim)
```
and, postprocess:
```{r genotype_sims, cache=TRUE, depends="setup_sims"}
rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
loci <- list(0.5+c(rev(rvals[-1]),rvals))
genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
freqs <- apply(genotypes,1,function(Z){
            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
        })
```

@@

Here's the PDE solution:
```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
zone.age <- sim$pars["n.gen"]
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                                                s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
pde.clines <- lapply( pde.solns, cline_from_soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
for (k in floor(seq(2,length(rvals),length.out=3))) {
    image(pde.solns[[k]],legend=TRUE,main=sprintf("r=%f",rvals[k]))
}
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)
    lines(1-thegrid$x.mid,match.clines[,k],col='green',lwd=3)
}
```

>>>>

Let's solve this numerically.
```{r solve_pde, cache=TRUE, dependson="solve_pde_setup"}
r <- 0.1
s <- 0.1
sigma <- 5
zone.age <- 200  # this is tau
tt <- seq(0,zone.age,length.out=100)
pde.soln <- solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,dependson=c("solve_pde_setup","solve_pde")}
pde.cline <- cline_from_soln(pde.soln,grid=thegrid)
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r),yaxt='n')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
# same thing, plotted with lines
clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
```
**Note:** the range should be wide enough these are *going to zero*
at the ends, or else we have to worry about the boundary conditions.

@@

We'll setup, run, and summarize simulations
as described in [this file](../sims/run-sims.html).
```{r setup_sims, cache=TRUE}
source("../sims/sim-fns.R")
zone.age    = 50
ndemes 		= 500
deme.size 	= 5
ninds 		= ndemes*deme.size
qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s, pos=c(0.5)))
# spatial locations
xx <- (1:ndemes)-0.5-ndemes/2
# run the simulation
sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=sigma,quiet=TRUE)
# post-process
sim.sums <- spBreaks(sim)
```
and, postprocess:
```{r genotype_sims, cache=TRUE, dependson="setup_sims"}
rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
loci <- list(0.5+c(rev(rvals[-1]),-rvals))
genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
freqs <- apply(genotypes,1,function(Z){
            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
        })
```

@@

Here's the PDE solution:
```{r match_pde,cache=TRUE,dependson="solve_pde_setup"}
zone.age <- sim$pars["n.gen"]
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve_pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                                                s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
pde.clines <- lapply( pde.solns, cline_from_soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
for (k in floor(seq(2,length(rvals),length.out=3))) {
    image(pde.solns[[k]],legend=TRUE,main=sprintf("r=%f",rvals[k]))
}
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    matplot(xx,freqs[,which(abs(rvals[k]-loci[[1]]+0.5)<1e-9)],type='l',
            xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)
    lines(1-thegrid$x.mid,match.clines[,k],col='green',lwd=3)
}
```
